Comb associative operads

### 2018-09-01 CAs_operads.tex
Harmonisation presentation/orientation

In this section, we define comb associative operads and we show that the
set of such operads admits a lattice structure, isomorphic to the lattice
division for nonnegative integers. We relate this lattice to the
lattice of linear magmatic quotients. We also provide a finite convergent
orientation for the comb associative operad corresponding to $3$.
\medbreak

@@

\Todo{Christophe : je vous propose cette démo}
\begin{proof}
    We consider the rewrite rule $\LComb{\gamma} \Rew \RComb{\gamma}$ 
    on $\Mag$. First, we show that 
    \begin{equation} \label{eq:rewriting_invariant}
        \Tfr \RewContext \Tfr' 
        \mbox{ implies that }
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}.
    \end{equation}
    Indeed for all $\Tfr \in \Mag$, there exist binary trees 
    $\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)}$ such that
    \begin{math}
      \Tfr = 
      \left[\, |,\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)} \right].
    \end{math}
    Thus if $\Tfr \RewContext \Tfr'$, then one of the 
    two following cases occurs :
    \begin{enumerate}[label={(\it\roman*)}]
        \item there exist $\Tfr_i'$ such that
            \begin{math}
              \Tfr' = \LComb{\LRank\left(\Tfr\right)} \circ 
              \left[\, | ,\Tfr_1, \dots, \Tfr_i',\dots, 
                    \Tfr_{\LRank\left(\Tfr\right)}
              \right] \mbox{ so } 
              \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right);
            \end{math}
        \item there exist $i$ such that
            \begin{math} 
              \Tfr' = \LComb{\LRank\left(\Tfr\right)-\bar{\gamma}} \circ
              \left[ | ,\Tfr_1, \dots, 
                 \RComb{\bar{\gamma}} \circ 
                 [\Tfr_{i},\dots, \Tfr_{i+\bar{\gamma}}]
                 ,\dots, \Tfr_{\LRank\left(\Tfr\right)}
              \right]\\
              \mbox{ so }
              \LRank\left(\Tfr'\right) = 
              \LRank\left(\Tfr\right) - \bar{\gamma} = 
              \LRank\left(\Tfr\right) \pmod{\bar{\gamma}}
            \end{math}.
    \end{enumerate} 
    This concludes the proof of~\ref{eq:rewriting_invariant}.  
    \smallbreak

    We deduce from~\ref{eq:rewriting_invariant} that the equivalence 
    relation $r$ defined by 
    \begin{equation}
        r(\Tfr, \Tfr') \mbox{ if and only if } 
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}
    \end{equation}
    contains the relation $\RewContext$ and thus its transitive 
    and symmetric closure, which is $\CongrCAs{\gamma}$.
\end{proof}

\begin{proof}
  We consider the rewrite rule $\LComb{\gamma} \Rew \RComb{\gamma}$ on
  $\Mag$.
    \smallbreak

    First, we show that $\Tfr \RewContext \Tfr'$ implies that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by $\bar{\gamma}$. \Todo{à faire.}
    \smallbreak

    Assume that $\Tfr \RewContextRT \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in [k - 1]$, we have
    $\Tfr_i\RewContext \Tfr_{i + 1}$. The integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\bar{\gamma}$ from the first part of the proof, so
    that $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by~$\bar{\gamma}$.
    \smallbreak

    Assume that $\Tfr \RewContextRST \Tfr'$, so that there exist binary trees
    $\Tfr_1$, \dots, $\Tfr_k$ such that $\Tfr_1 = \Tfr$,
    $\Tfr_k = \Tfr'$ and for every $i \in [k - 1]$, we have
    $\Tfr_i \RewContextRT \Tfr_{i + 1}$ or $\Tfr_{i + 1} \RewContextRT
    \Tfr_i$. From the second part of the proof, the integer
    \begin{math}
        \LRank\left(\Tfr_{i + 1}\right)
        -
        \LRank\left(\Tfr_i\right)
    \end{math}
    is divisible by $\bar{\gamma}$, so that
    $\LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)$ is
    divisible by~$\bar{\gamma}$.
    \smallbreak

    The congruence $\CongrCAs{\gamma}$ is the equivalence relation
    induced by the rewrite relation $\RewContext$ induced by the rewrite
    rule $\LComb{\gamma} \Rew \RComb{\gamma}$, that is we have $\Tfr
    \CongrCAs{\gamma} \Tfr'$ if and only if $\Tfr \RewContextRST \Tfr'$.
    Hence, from the third part of the proof, $\Tfr \CongrCAs{\gamma}
    \Tfr'$ implies that
    \begin{math}
        \LRank\left(\Tfr'\right) - \LRank\left(\Tfr\right)
    \end{math}
    is divisible by~$\bar{\gamma}$.
\end{proof}
\medbreak

@@

\begin{Theorem} \label{thm:convergent_rewrite_rule_CAs_3}
    The set $\Rew$ of rewrite rules containing~\eqref{equ:rew_1},
    \eqref{equ:rew_2}, \eqref{equ:rew_3}---\eqref{equ:rew_11} is a
    convergent orientation of the congruence $\CongrCAs{3}$
    of~$\CAs{3}$.
\end{Theorem}
\begin{proof}
    Let us show that the rewrite relation $\RewContext$ induced by
    $\Rew$ is convergent. First, for every relation $\Tfr \Rew \Tfr'$,
    we have $\Tfr > \Tfr'$. Therefore, by
    Lemma~\ref{lem:prefix_word_termination}, $\RewContext$ is
    terminating. Moreover, the greatest degree of a tree appearing in
    $\Rew$ is~$7$ so that, from Lemma~\ref{lem:degree_confluence}, to
    show that $\RewContext$ is convergent, it is enough to prove that
    each tree of degree at most $13$ admits exactly one normal form.
    Equivalently, this amounts to show that the number of normal forms
    of trees of arity $n$ is equal to $\#\CAs{3}(n)$. By computer
    exploration, we get the same sequence
    \begin{equation} \label{equ:dimensions_CAs_3}
        1, 1, 2, 4, 8, 14, 20, 19, 16, 14, 14, 15, 16, 17
    \end{equation}
    for $\#\CAs{3}(n)$ and for the numbers of normal forms of arity $n$,
    when $ 1 \leq n \leq 14$. Hence, we get our following main result.
\end{proof}
\medbreak

>>>>

In this section, we define comb associative operads and we show that the
set of such operads admits a lattice structure, isomorphic to the lattice
division for nonnegative integers. We relate this lattice to the
lattice of linear magmatic quotients. We also provide a finite convergent
presentation of the comb associative operad corresponding to $3$.
\medbreak

@@

\begin{proof}
    We consider the rewrite rule $\LComb{\gamma} \Rew \RComb{\gamma}$ 
    on $\Mag$. First, we show that 
    \begin{equation} \label{eq:rewriting_invariant}
        \Tfr \RewContext \Tfr' 
        \mbox{ implies }
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}.
    \end{equation}
    Indeed, for all $\Tfr \in \Mag$, let
    $\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)}$ be the binary trees
    such that
    \begin{math}
      \Tfr = 
      \left[\, |,\Tfr_1, \dots, \Tfr_{\LRank\left(\Tfr\right)} \right].
    \end{math}
    Thus, if $\Tfr \RewContext \Tfr'$, then one of the following two
    cases occurs:
    \begin{enumerate}[label={(\it\roman*)}]
        \item there exists $\Tfr_i'$ such that
            \begin{math}
              \Tfr' = \LComb{\LRank\left(\Tfr\right)} \circ 
              \left[\, | ,\Tfr_1, \dots, \Tfr_i',\dots, 
                    \Tfr_{\LRank\left(\Tfr\right)}
              \right] \mbox{, so that } 
              \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right);
            \end{math}
        \item there exists $i$ such that
            \begin{math} 
              \Tfr' = \LComb{\LRank\left(\Tfr\right)-\bar{\gamma}} \circ
              \left[ | ,\Tfr_1, \dots, 
                 \RComb{\bar{\gamma}} \circ 
                 [\Tfr_{i},\dots, \Tfr_{i+\bar{\gamma}}]
                 ,\dots, \Tfr_{\LRank\left(\Tfr\right)}
              \right]
              \mbox{, so that }\\
              \LRank\left(\Tfr'\right) = 
              \LRank\left(\Tfr\right) - \bar{\gamma} = 
              \LRank\left(\Tfr\right) \pmod{\bar{\gamma}}
            \end{math}.
    \end{enumerate} 
    This concludes the proof of~\eqref{eq:rewriting_invariant}.  
    \smallbreak

    We deduce from~\eqref{eq:rewriting_invariant} that the equivalence 
    relation $r$ defined by 
    \begin{equation}
        r(\Tfr, \Tfr') \mbox{ if and only if } 
        \LRank\left(\Tfr'\right) = \LRank\left(\Tfr\right) 
        \, \pmod{\bar{\gamma}}
    \end{equation}
    contains the relation $\RewContext$, so that it contains its
    transitive symmetric closure. The latter is $\CongrCAs{\gamma}$,
    which concludes the proof of Lemma~\ref{lem:left_rank_and_CongrCAs}.
\end{proof}
\medbreak

@@

\begin{Theorem} \label{thm:convergent_rewrite_rule_CAs_3}
    The set $\Rew$ of rewrite rules containing~\eqref{equ:rew_1},
    \eqref{equ:rew_2}, \eqref{equ:rew_3}---\eqref{equ:rew_11} is a
    convergent presentation of~$\CAs{3}$.
\end{Theorem}
\begin{proof}
    Let us show that the rewrite relation $\RewContext$ induced by
    $\Rew$ is convergent. First, for every relation $\Tfr \Rew \Tfr'$,
    we have $\Tfr > \Tfr'$. Therefore, by
    Lemma~\ref{lem:prefix_word_termination}, $\RewContext$ is
    terminating. Moreover, the greatest degree of a tree appearing in
    $\Rew$ is~$7$ so that, from Lemma~\ref{lem:degree_confluence}, to
    show that $\RewContext$ is convergent, it is enough to prove that
    each tree of degree at most $13$ admits exactly one normal form.
    Equivalently, this amounts to show that the number of normal forms
    of trees of arity $n$ is equal to $\#\CAs{3}(n)$. By computer
    exploration, we get the same sequence
    \begin{equation} \label{equ:dimensions_CAs_3}
        1, 1, 2, 4, 8, 14, 20, 19, 16, 14, 14, 15, 16, 17
    \end{equation}
    for $\#\CAs{3}(n)$ and for the numbers of normal forms of arity $n$,
    when $ 1 \leq n \leq 14$. Hence, we get our following main result.
\end{proof}
\medbreak

### 2018-09-01 Conclusion.tex
Harmonisation presentation/orientation

A first research direction consists in finding convergent orientations
of presentations for (all or most of) the $\CAs{\gamma}$ operads in
order to describe algebraic and combinatorial properties of these
operads (as describing explicit bases, computing Hilbert
series, and providing combinatorial realizations). This has been
done only in the case $\gamma = 3$. For some other cases, we only have
conjectural and experimental data (see Section~\ref{?}).
\Todo{S: compléter avec la bonne partie une fois qu'on aura ajouté ces
données}.
\medbreak

Following ideas existing for word rewriting theory~\cite{GGM15}, a
second axis consists in allowing new generators in the orientations of
the operad congruence $\CongrCAs{\gamma}$ in order to obtain finite
convergent orientations for their presentations when $\gamma \geq 4$.
Indeed, the Buchberger semi-algorithm works by adding rewrite rules to a
set of rewrite rules to obtain a convergent rewrite system. An orthogonal
procedure consists rather in adding new generators (new labels for
internal nodes in the trees) in order to obtain convergent rewrite
systems. More generally, we also would like to use these ideas for other
magmatic quotients, such as the operad $\Mag^{\{3, 4\}}$ that we did not
study entirely in Section~\ref{sec:MAg_3}.
\medbreak

A third axis consists in studying if the completion of orientations of
presentations of quotients of the magmatic operad maintains links with
the lattice structure introduced in Section~\ref{sec:Magmatic_operads}.
More precisely, assuming that we have completed the orientations of
the presentations of the quotients $\Oca_1$ and $\Oca_2$ of $\Mag$,
as well as the one of the lower-bound $\Oca_2 \InfQMag \Oca_2$, the
question consists in designing an algorithm for computing a completion
of an orientation of the presentation of the upper-bound
$\Oca_1 \SupQMag \Oca_2$. Of course, the same question also makes sense
for the lattice of comb associative operads introduced in
Section~\ref{sec:CAs_d}.
\medbreak

@@

\Todo{S: attention, on emploie souvent l'expression "presentation" et
"completion of presentation" mais on a jamais vraiment défini ce qu'est
une presentation et ce qu'est une orientation d'une complétion. C'est
facile, mais il faudrait le faire (préliminaires). Il faudrait d'ailleurs
utiliser plutôt l'expression "convergent orientation of a presentation"
plutôt que "convergent orientation".}

>>>>

A first research direction consists in finding convergent presentations
for (all or most of) the $\CAs{\gamma}$ operads in
order to describe algebraic and combinatorial properties of these
operads (as describing explicit bases, computing Hilbert
series, and providing combinatorial realizations). This has been
done only in the case $\gamma = 3$. For some other cases, we only have
conjectural and experimental data (see Section~\ref{?}).
\Todo{S: compléter avec la bonne partie une fois qu'on aura ajouté ces
données}.
\medbreak

Following ideas existing for word rewriting theory~\cite{GGM15}, a
second axis consists in allowing new generators for the operads
$\CAs{\gamma}$ in order to obtain finite convergent presentations
when $\gamma \geq 4$. Indeed, the Buchberger semi-algorithm works by
adding rewrite rules to a set of rewrite rules to obtain a convergent
rewrite system. An orthogonal procedure consists rather in adding new
generators (new labels for internal nodes in the trees) in order to
obtain convergent rewrite systems. More generally, we also would like to
use these ideas for other magmatic quotients, such as the operad
$\Mag^{\{3, 4\}}$ that we did not study entirely in
Section~\ref{sec:MAg_3}.
\medbreak

A third axis consists in studying if the completion of presentations of
quotients of the magmatic operad maintains links with
the lattice structure introduced in Section~\ref{sec:Magmatic_operads}.
More precisely, assuming that we have completed the presentations of
the quotients $\Oca_1$ and $\Oca_2$ of $\Mag$,
as well as the one of the lower-bound $\Oca_2 \InfQMag \Oca_2$, the
question consists in designing an algorithm for computing a completion
of a presentation of the upper-bound
$\Oca_1 \SupQMag \Oca_2$. Of course, the same question also makes sense
for the lattice of comb associative operads introduced in
Section~\ref{sec:CAs_d}.
\medbreak

@@

### 2018-09-01 Operads_rewrite.tex
Maj partie CAs
Harmonisation presentation/orientation

\begin{Lemma} \label{lem:normal_forms_avoiding}
    Let $\Rew$ be a set of rewrite rules on $\Mag$ and $\RewContext$ be
    the rewrite relation induced by $\Rew$. Then,
    %if $\RewContext$ is terminating,
    $\NormalForms_{\RewContext}$ is the set of all the binary trees that
    avoid all the trees appearing as left members of~$\Rew$.
\end{Lemma}
\begin{proof}
    Assume first that $\Tfr$ is a binary tree avoiding all the trees
    appearing as left members of~$\Rew$. Then, due to the
    definition~\eqref{equ:rewrite_relation_induced} of $\RewContext$,
    $\Tfr$ is not rewritable by $\RewContext$. Hence, $\Tfr$ is a normal
    form for $\RewContext$. Conversely, assume that
    $\Tfr \in \NormalForms_{\RewContext}$. In this case, by definition
    of a normal form, $\Tfr$ is not rewritable by $\RewContext$, so that
    $\Tfr$ does not admit any occurrence of a tree appearing as a left
    member of~$\Rew$.
\end{proof}
\medbreak

@@

Let us now go back on operads. Let $\Congr$ be an operad congruence of
$\Mag$. If $\Tfr$ is a binary tree, we denote by $[\Tfr]_{\Congr}$ the
$\Congr$-equivalence class of $\Tfr$. By definition, $[\Tfr]_{\Congr}$
is an element of the quotient operad
\begin{equation} \label{equ:quotient_operad_mag}
    \Oca := \Mag/_{\Congr}.
\end{equation}
A set of rewrite rules $\Rew$ is an \Def{orientation} of $\Congr$ if
$\RewContextRST$ and $\Congr$ are equal as binary relations, where
$\RewContext$ is the rewrite relation induced by $\Rew$. Moreover,
$\Rew$ is a \Def{convergent} (resp. \Def{terminating}, \Def{confluent})
orientation of $\Congr$ if $\RewContext$ is convergent (resp.
terminating, confluent).
\medbreak

>>>>

\begin{Lemma} \label{lem:normal_forms_avoiding}
    Let $\Rew$ be a set of rewrite rules on $\Mag$ and $\RewContext$ be
    the rewrite relation induced by $\Rew$. Then,
    $\NormalForms_{\RewContext}$ is the set of all the binary trees that
    avoid all the trees appearing as left members of~$\Rew$.
\end{Lemma}
\begin{proof}
    Assume first that $\Tfr$ is a binary tree avoiding all the trees
    appearing as left members of~$\Rew$. Then, due to the
    definition~\eqref{equ:rewrite_relation_induced} of $\RewContext$,
    $\Tfr$ is not rewritable by $\RewContext$. Hence, $\Tfr$ is a normal
    form for $\RewContext$. Conversely, assume that
    $\Tfr \in \NormalForms_{\RewContext}$. In this case, by definition
    of a normal form, $\Tfr$ is not rewritable by $\RewContext$, so that
    $\Tfr$ does not admit any occurrence of a tree appearing as a left
    member of~$\Rew$.
\end{proof}
\medbreak

@@

Let us now go back on operads. Let $\Congr$ be an operad congruence of
$\Mag$. If $\Tfr$ is a binary tree, we denote by $[\Tfr]_{\Congr}$ the
$\Congr$-equivalence class of $\Tfr$. By definition, $[\Tfr]_{\Congr}$
is an element of the quotient operad
\begin{equation} \label{equ:quotient_operad_mag}
    \Oca := \Mag/_{\Congr}.
\end{equation}
A set of rewrite rules $\Rew$ is an \Def{orientation} of $\Congr$ if
$\RewContextRST$ and $\Congr$ are equal as binary relations, where
$\RewContext$ is the rewrite relation induced by $\Rew$. Moreover,
$\Rew$ is a \Def{convergent} (resp. \Def{terminating}, \Def{confluent})
orientation of $\Congr$ if $\RewContext$ is convergent (resp.
terminating, confluent). In the sequel, we use the terminology of a
\Def{presentation} of a magmatic operad $\Oca$: it is the data of an
operad congruence $\Congr$ such that $\Oca$ is isomorphic to 
\begin{math}
  \Mag/_{\Congr}.
\end{math}
For simplicity, we say that a convergent (resp. terminating, confluent)
orientation of $\Congr$ is a \Def{convergent} (resp. \Def{terminating},
\Def{confluent}) \Def{presentation} of $\Oca$.
\medbreak
